---
title: "Analyze_Zika_Dataset"
author: "Sunil Iyengar"
date: "August 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Cleaning and dataset manupulation 
```{r clean}
dszika <- read.csv("C://Temp//zika//cdc_zika.csv",na.strings = c(""))
dszika$value <- as.numeric(dszika$value)
dszika$report_date <- as.Date(dszika$report_date)

#Extract Country
dszika$Country <- substr(dszika$location,start = 1, stop = regexpr(pattern = "[[:alnum:]]-",text = dszika$location))
dszika$Country <- factor(dszika$Country)

#Extract State
temp <- substr(dszika$location,start = regexpr(pattern = "-[[:alnum:]]-*",text = dszika$location), stop = nchar(as.character(dszika$location)))
temp <- substr(temp, start = regexpr(pattern = "-*",text = temp) + 1, stop = nchar(temp))
dszika$State <- substr(temp, start = 1, stop = regexpr(pattern = "[[:alnum:]]-",text = temp))
dszika$State <- factor(dszika$State)

#Extract Year from Report Date
dszika$Year <- substr(dszika$report_date,start = 1, stop = regexpr(pattern = "[[:alnum:]]-",text = dszika$report_date))
dszika$Year <- factor(dszika$Year)

#Extract Month from Report Date
dszika$Month <- substr(x = dszika$report_date, start = regexpr(pattern = "-+", text = dszika$report_date) + 1, stop=regexpr(pattern = "-+", text = dszika$report_date) + 2)
dszika$Month <- factor(dszika$Month)

#Remove columns that have no value
dszika$time_period <- NULL
dszika$time_period_type <- NULL

#Create dataset with no NA in the Report_Date field
dszika1 <- dszika[which(complete.cases(dszika$report_date)),]

```

## Summary of the resultant dataset after manipulation

```{r summary}
summary(dszika1)
str(dszika1)

```

##Plots and analysis

<p>Plot of Number of cases reported across various dates</p>
```{r plot1, echo=FALSE}
library(ggplot2)
dsZikaObjBase <- ggplot(data=dszika1, aes(x=format(report_date,"%b/%Y"), color=Country))
dsZikaObjBase + geom_bar(aes(fill=Country),position = "dodge") + xlab("Reported Date") + ylab("No. of Cases Reported")
```
It is very clear from the above plot that Colombia has the highest reported cases. In fact, 81% of cases reported are from Colombia</b>

```{r }
#Percentage of Colombia cases 
paste(round((nrow(dszika1[dszika1$Country=="Colombia",]) / nrow(dszika1)) * 100,2),"%")
```

<p>Let us see the number of cases report across different months</p>

```{r, echo=FALSE}
dsZikaObjBase + geom_density(aes(fill=Country),position = "stack") +  xlab("Reported Date") + ylab("No. of Cases Reported")
```

<p>Wow! April has the highest reported cases</p>
<p>Okay, now lets analyze Colmbia, since it has the bulk of cases reported. <BR>
What percentage are confirmed cases?</p>

```{r}
#Percentage of Colombia cases - Confirmed cases vs total cases 
paste(round(nrow(dszika1[dszika1$Country=="Colombia" & (dszika1$data_field_code == "CO0001" | dszika1$data_field_code == "CO0002"),]) / nrow(dszika1[dszika1$Country=="Colombia",]) * 100,2), "%")
```
<p>Interestingly almost <b>50% are confirmed cases</b>. The remaining <b>50% are suspected</b> cases</p>

<p>Now let us visualize break up as per case type</p>

```{r plot4, echo=FALSE}
#Create a new data set for Colombia and do some more analysis
dszikaColombia <- dszika1[dszika1$Country=="Colombia",]

#Case type wise breakup for Colombia
ggplot(data=dszikaColombia, aes(x=format(report_date,"%b/%Y"), color=data_field)) + geom_bar(aes(fill=data_field),position = "dodge") + xlab("Reported Date") + ylab("No. of Cases Reported")
```

<H3>That is it for now. Deeper analysis coming up! Check back for updates.</H3>
